\bf Model Visualization Techniques for a Social Network Model

### 2018-07-17 paper-child.Rnw
ugh forgot to :arrow_up: last night. :anger:

\section{Introduction}

@@

\section{Networks and their Visualizations} \label{sec:netintro}

@@

\subsection{Introduction to Network Structures}

@@

\subsection{Visualizing Network Data} \label{sec:netvizintro}

@@

\section{Defining SAOMs} \label{sec:saoms}

@@

\subsection{Definitions, Terminology, and Notation}

@@

The CTMC process $X(T)$ is a series of single tie changes that happen according to some pre-defined rate function (Section \ref{sec:saomratefn}), where one actor at a time is selected and it adds or removes one outgoing tie, or does nothing. Once an actor is chosen at random according to the rate function, it is ``given" the chance to change a tie, and it tries to maximize its utility function (Section \ref{sec:saomobjfn}) based on the current and possible future states of the network.

\subsubsection{The Rate Function}\label{sec:saomratefn}

In general, each actor $i$ in the network $x$ may change its ties, $x_{ij}$, to other nodes $j \neq i$ according to the SAOM's  \emph{rate function}, $\rho(x, \mathbf{z}, \boldsymbol{\alpha})$, where $\boldsymbol{\alpha}$ are the parameters in the function $\rho$, and $x$ is the current network state, with covariates of interest $\mathbf{z}$. With this general formulation, it is possible that each node will have a different rate of change. We assume a simple rate function, $\rho(x, \mathbf{z}, \boldsymbol{\alpha}) = \alpha_m$ that is constant across all actors between time $t_m$ and $t_{m+1}$. With this simple rate function, we have just one set of rate parameters in the overall model. In this simple model, the rate parameters dictate how often an actor $i$ ``gets an opportunity" to change one of its ties, $x_{ij}$, for $j \in \{1, \dots, n\}$ in the time period from $t_{m}$ to $t_{m+1}$ for $m = 1, \dots, M-1$ (Note that when $j = i$, no change is made.) We also assume that the actors $i$ are conditionally independent given their ties, $x_{i1}, \dots, x_{in}$ at the current network state. Let $\tau(i|x,m) $ be the wait time until actor $i$ gets to the opportunity to change its current set of ties. For any time point, $T$, where $t_{m} \leq T < t_{m+1}$, the waiting time to the next change made by actor $i$ is exponentially distributed with expected value $\alpha_m^{-1}$. Then, we express this in Equation~\ref{eq:c1}.

@@

\subsubsection{The Objective Function}\label{sec:saomobjfn}

Because of the conditional independence assumptions given in Equation~\ref{eq:c1}, we can consider the objective function for each node separately, as only one tie from one node is changing at a time in the CTMC. The node $i$, which is the node that is given the opportunity to change at the current time point, is called the \textit{ego} node. It has the potential to interact with all other nodes in the network, $j \neq i$. These nodes $j$, are referred to as \textit{alter} nodes, and do not change any of their ties when $i$ may change. For the ego node $i$ in the current network state $x$, its \emph{objective function}, which it tries to maximize, is written as
\begin{equation} \label{eq:of1}
f_i(\boldsymbol{\beta}, x) = \sum_k \beta_k s_{ik}(x, \mathbf{Z}),
\end{equation}
for $x \in \mathcal{X}$, the space of all possible directed networks with the $n$ nodes, and $\mathbf{Z}$, the matrix of covariates. The vector $\boldsymbol{\beta}$ contains the parameters of the model with corresponding network and covariate statistics, $s_{ik}(x, \mathbf{Z})$, for $k = 1,\dots, K$. To increase the value of $f_i$, $i$ will destroy one tie where $x_{ij} = 1$, create one tie where $x_{ij}=0$, or make no change if any changes result in lower values of $f_i$.

@@

\subsection{Fitting Models to Data}

@@

\subsection{Example Data}

<<adjmatviz, fig.align='center', fig.cap="A visualization of the adjacency matrices of the three waves of network observations in the ``Teenage Friends and Lifestyle Study'' data. The subset we will be using is outlined in red.", results='hide',fig.show='asis', fig.width=8, fig.height=3>>=
source("code/00e_small_friends.R")
view_rs2$wave <- factor(view_rs2$wave)
levels(view_rs2$wave) <- c("Wave 1", "Wave 2", "Wave 3")
ggplot(data= view_rs2, aes(x = y, y=-x)) +
  geom_tile(aes(fill = as.factor(value))) +
  scale_fill_manual("i likes j", labels=c("no", "yes"),
                    values = c('white', 'black')) +
  geom_rect(data= NULL, inherit.aes = FALSE, color = 'red',
            aes(xmin = 19.5, xmax = 35.5, ymax = -19.5, ymin = -35.5, fill =NA)) +
  facet_wrap(~wave) + theme_bw() +
  labs(x = "node j (alter)", y = "node i (ego)") +
  ThemeNoNet +
  theme(aspect.ratio=1, axis.text = element_blank(), axis.ticks = element_blank(), panel.grid = element_blank())
@

<<alldataviz, fig.cap="The smaller friendship network data we will be modelling throughout the paper. The nodes change color as their drinking behavior changes, and the network becomes less densely connected over time. These are the mechanisms we hope to capture with SAOMs.", out.width='.99\\textwidth', fig.height=3>>=
actual1 <- merge(data.frame(as.edgelist(as.network(fd2.w1))),
                 data.frame(id = 1:16, drink = drink2[,1]),
                 by.x = "X1", by.y = "id", all = T)
actual1$Wave <- 1
actual2 <- merge(data.frame(as.edgelist(as.network(fd2.w2))),
                 data.frame(id = 1:16,drink = drink2[,2]),
                 by.x = "X1", by.y = "id", all = T)
actual2$Wave <- 2
actual3 <- merge(data.frame(as.edgelist(as.network(fd2.w3))),
                 data.frame(id = 1:16,drink = drink2[,3]),
                 by.x = "X1", by.y = "id", all = T)
actual3$Wave <- 3
alldat <- rbind(actual1, actual2, actual3)
#alldat$X1 <- paste0("V", alldat$X1)
alldat$X1 <- as.factor(alldat$X1)
#alldat$X2 <- ifelse(!is.na(alldat$X2), paste0("V", alldat$X2), NA)
alldat$X2 <- as.factor(alldat$X2)
#write_csv(alldat, "smallfriends4Geomnet.csv")
alldataviz <- ggplot(data = alldat,
       aes(from_id = X1, to_id = X2, color = as.ordered(drink), group = as.ordered(drink))) +
  geom_net(fiteach = FALSE, directed=TRUE, layout.alg = 'fruchtermanreingold', size = 5, arrowsize = .25, arrowgap = .05, ecolour = 'grey40', labelon=T, labelcolour = 'black', fontsize = 4, vjust = .5, linewidth=0.5) +
  scale_color_brewer(palette = "YlOrRd", name = "Drinking\nBehavior", labels = c("never", "1-2 times per year", "once per month", "once per week")) +
  ThemeNet +
  facet_wrap(~Wave, labeller = "label_both") +
  theme(legend.position = 'bottom',
        panel.background = element_rect(fill = NA, color = "grey30"),
        legend.text = element_text(size = 12, family = "Times New Roman"),
        legend.title = element_text(size = 12, family = "Times New Roman"))
alldataviz
@

To guide our visual exploration of SAOMs, we use two example data sources. The first is a subset of the 50 actor dataset from the ``Teenage Friends and Lifestyle Study" that is provided on the \texttt{RSiena} webpage. These data come from \citet{friendsdata}, and we chose to only work with a subset of the data to make node-link diagrams easier to see and to make any changes in the network more noticeable. %To determine which subset to select, we visualized all waves of the full network using the adjacency matrix visualization approach, which we show in Figure~\ref{fig:adjmatviz}. This adjacency matrix visualization is different from the one in Figure~\ref{fig:toyplots} because it does not show the node IDs on the axes. The network in Figure~\ref{fig:adjmatviz} is much larger than our toy data example, so we remove the node labels to remove clutter. In all three adjacency matrices, the ego and the alter nodes are ordered by node ID, 1-50, which were determined arbitrarily by the relationships in wave 1. 
In Figure~\ref{fig:adjmatviz}, we show the adjacency matrix visualization of all three observations of this network, with the subset we selected for analysis outlined in red. This subset contains 16 actors, their ties and their drinking behavior for three timepoints. The node-link visualization of the data is provided in Figure~\ref{fig:alldataviz}. For model fitting, we condition on wave 1 and estimate the parameters of the models for the second and third waves. The actor-level categorical covariate, drinking behavior, has four values: (1) does not drink, (2) drinks once or twice a year, (3) drinks once a month, and (4) drinks once a week. In Figure~\ref{fig:alldataviz}, the nodes are colored according to the drinking behavior of that student. Over time, we can see that the students tend to drink more and become increasingly isolated into smaller groups. An analysis of this type of data with a SAOM should capture these dynamics in a way that allows the researcher to draw conclusions about the nature of the network and behavioral forces at play.

@@

We use these two example newtoks to demonstrate our model visualization methods for SAOMs in section~\ref{sec:ModelVis}

\section{Model Visualizations} \label{sec:ModelVis}

\subsection{The Models}\label{sec:models}

@@

\subsection{View the model in the data space}\label{sec:minds}

@@

<<summnetM1, message=FALSE, fig.cap='On the left, the first wave of observed data that is conditioned on in the model. On the right, the second wave of observed data. In the middle, a summary network from the first model fit to the data. This summary network represents 1,000 simulations of wave 2 using the values from the simple fitted model M1.',results='hide', fig.height=2.5>>=
friendshipData <- array(c(fd2.w1, fd2.w2, fd2.w3), dim = c(16, 16, 3))
friendshipData <- array(c(fd2.w1, fd2.w2, fd2.w3), dim = c(16, 16, 3))
drink1 <- drink[20:35,1]
drink2 <- drink[20:35,2]
drink3 <- drink[20:35,3]
friendship <- sienaDependent(friendshipData)
alcohol <- varCovar(cbind(drink1,drink2, drink3))
mydata <- sienaDataCreate(friendship, alcohol)
M1eff <- getEffects(mydata)
M2eff <- includeEffects(M1eff, jumpXTransTrip, interaction1 = "alcohol")
M3eff <- includeEffects(M1eff, nbrDist2twice)
simu2 <- read.csv("data/simulation-1000-M1-M2-M3.csv")
means <- simu2 %>% group_by(Model, parameter) %>% summarize(
  means = mean(estimate)
)
M1parms <- (means %>% filter(Model == "M1"))$means
M2parms <- (means %>% filter(Model == "M2"))$means
M3parms <- (means %>% filter(Model == "M3"))$means
set.seed(4231352)
M1sims1000 <- netvizinf::saom_simulate(dat = mydata,
                            struct = M1eff,
                       parms = as.numeric(M1parms),
                       N = 1000)
M2sims1000 <- netvizinf::saom_simulate(dat = mydata,
                            struct = M2eff,
                       parms = as.numeric(M2parms),
                       N = 1000)
M3sims1000 <- netvizinf::saom_simulate(dat = mydata,
                            struct = M3eff,
                       parms = as.numeric(M3parms),
                       N = 1000)
M1simsdf <- netvizinf::sims_to_df(M1sims1000)
M2simsdf <- netvizinf::sims_to_df(M2sims1000)
M3simsdf <- netvizinf::sims_to_df(M3sims1000)
M1avgW2 <- M1simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))
M2avgW2 <- M2simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))
M3avgW2 <- M3simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))

@@

For network data, one way we view an ``average" network is through a sory of summary network drawn using the traditional node-link diagram. In Figure~\ref{fig:summnetM1}, we show an \textit{average network} created with 1,000 simulations of the second wave of the network from Model 1. To make this average network, we first simulated 1,000 wave 2 and wave 3 observations of our small friendship example data from model M1, for which parameters had previously been estimated. We then combine the 1,000 simulated instances of wave 2, and count up the number of times each edge appears in a network simulation. Then, we combine these 1,000 networks into a single network with edgeweight equal to the proportion of time that edge appears in our 1,000 simulations. This weighted edgelist is the network we draw in Figure~\ref{fig:summnetM1}. An edge is only drawn in the average network if it appears in at least 51 of the 1000 simulations, with edges that appear with greater frequency emphasized by thicker linewidths and a darker color. On either side of the average network in Figure~\ref{fig:summnetM1}, we show the actual data, wave 1 on the left, and wave 2 on the right. We can see that the structure of the average network is much more similar to the first wave than to the second wave. However, the simulations are supposed to represent the second wave of data on the right of Figure~\ref{fig:summnetM1}. This is an indication that the simple model, M1, is doing a very poor job of capturing the change mechanism from the first to the second wave of observation. The average network can thus be used to help determine model goodness-of-fit. Because the the average network looks more like the first wave than the second wave, we can use the visualization in Figure~\ref{fig:summnetM1} as evidence of poor model fit.

@@

\begin{figure}
\centering
\includegraphics[width=.75\linewidth]{img/Model1LUSmallFriends.png}
\caption{\label{fig:lineup} A small lineup of node-link diagrams showing the second wave of our small friendship network among five networks simulated from model M1.}
\end{figure}

Another potential goodness-of-fit visualization that places the fitted model in the node and edge data space is a lineup like those proposed in \citet{buja:2009}. A \textit{lineup} ``asks the witness to identify the plot of the real data from among a set of decoys, the null plots, under the veil of ignorance" \citep[p. 4369]{buja:2009}. It can be thought of like a police lineup, where the ``suspect" is in a lineup among several innocent similar looking people, and if the witness picks the suspect out of the lineup, that is evidence the suspect is guilty.

In data and model visualization, the ``suspect" is a plot of the true data, while the ``filler" is composed of several plots of ``mock data", simulated from a hypothesized model. If the true data stands out among the simulated data, that is taken as evidence agains the hypothesised model. If, however, the true data is difficult to identify among the simulated data, that is taken as evidence in favor of the hypothesized model. 
% An application of the lineup protocal can be found in \citet{hofmann2012graphical}, where the authors examine, for instance, the differences between polar and cartesian coordinates for plotting categorical data, and density plots and box plots for determining distributional differences. They pose questions to experiment participants such as, ``Which plot is most different from the others?" or, ``In which plot is the blue group furthest to the right?"

The data visualizations examined in previous applications such as  \citet{hofmann2012graphical} are less complex than a node-link network diagram. Typically, the lineup viewers have to pick out the most different plot, which is hard for network data. As an example, consider the network lineup shown in Figure~\ref{fig:lineup}. In this small network lineup, the second wave of the small friendship network is shown among five simulated networks from model M1 using parameter values estimated from the data. It seems possible to argue for any one of the six plots in Figure~\ref{fig:lineup} as most different. So, we suggest guiding participants to look at the overall structure of the graphs to determine which has the most or least connected or complex structure.% The least complex plot, number six, has no triangles (3-cycles), while the most complex plot, number three, has three 3-cycles. The most complex plot, number three, is in fact the data. We have found that when shown lineups like that of Figure~\ref{fig:lineup}, online survey participants %where all the node-link diagrams shown are based on simulated data, %one observation from an ``alternative" model and the rest from a ``null" model, %pick up on the triangular shape of the 3-cycles most often. 
In a future paper, we will use the lineup protocol to help better understand SAOMs by allowing us to view the model in the data space. %Lineups place observations simulated from the model side-by-side with the data and let the viewer examining the differences. This can also help us determine the significant structural effects, if any, of the parameters in the model on observations simulated from the model. The triangular shape mentioned above becomes more prevalent when a transitive triplet parameter ($\beta_3$ in Section~\ref{sec:models}) is included in a model, indicating that it may have a significant effect on network structure. %, but it does not always cause triplets to form in the simulated data. This requires more investigation with larger data sets and more complex models, but is promising for the development of additional goodness-of-fit measures for network models.

% \st{
% \begin{itemize}
% \item WHat is the model? the network model (SAOM model is just one of many)
% \item What is the data? Edge data, node data, observed time points, unobserved timepoints (CTMC)
% \item Note: specifically dealing with dynamic networks observed at discrete time points.
% \end{itemize}
% What are some things that could go in here? need more data examples; not just the friends data.
% Harry Potter data??? I want to find my own data, too.}

\subsection{Visualizing collections of models}

@@

\subsubsection{Exploring the space of all possible models}\label{sec:allpossible}

@@

\subsubsection{Varying model settings}

@@

\subsubsection{Fitting the same model to different data}\label{sec:samemoddiffdat}

@@

\subsubsection{Fitting the same model to the same data}\label{sec:samemodsamedat}

@@

\subsection{Explore algorithms, not just end result}\label{sec:algorithms}

>>>>

\section{Introduction \& Background}

@@

\subsection{Networks and their Visualizations} \label{sec:netintro}

@@

\subsubsection{Introduction to Network Structures}

@@

\subsubsection{Visualizing Network Data} \label{sec:netvizintro}

@@

\subsection{Defining SAOMs} \label{sec:saoms}

@@

\subsubsection{Definitions, Terminology, and Notation}

@@

The CTMC process $X(T)$ is a series of single tie changes that happen according to a pre-defined rate function, where one actor at a time is selected and it adds or removes one outgoing tie, or does nothing. Once an actor is chosen at random according to the rate function, it is ``given" the chance to change a tie, and it tries to maximize its utility function based on the current and possible future states of the network.

\textbf{The Rate Function} In general, each actor $i$ in the network $x$ may change its ties, $x_{ij}$, to other nodes $j \neq i$ according to the SAOM's  \emph{rate function}, $\rho(x, \mathbf{z}, \boldsymbol{\alpha})$, where $\boldsymbol{\alpha}$ are the parameters in the function $\rho$, and $x$ is the current network state, with covariates of interest $\mathbf{z}$. With this general formulation, it is possible that each node will have a different rate of change. We assume a simple rate function, $\rho(x, \mathbf{z}, \boldsymbol{\alpha}) = \alpha_m$ that is constant across all actors between time $t_m$ and $t_{m+1}$. With this simple rate function, we have just one set of rate parameters in the overall model. In this simple model, the rate parameters dictate how often an actor $i$ ``gets an opportunity" to change one of its ties, $x_{ij}$, for $j \in \{1, \dots, n\}$ in the time period from $t_{m}$ to $t_{m+1}$ for $m = 1, \dots, M-1$ (Note that when $j = i$, no change is made.) We also assume that the actors $i$ are conditionally independent given their ties, $x_{i1}, \dots, x_{in}$ at the current network state. Let $\tau(i|x,m) $ be the wait time until actor $i$ gets to the opportunity to change its current set of ties. For any time point, $T$, where $t_{m} \leq T < t_{m+1}$, the waiting time to the next change made by actor $i$ is exponentially distributed with expected value $\alpha_m^{-1}$. Then, we express this in Equation~\ref{eq:c1}.

@@

\textbf{The Objective Function} Because of the conditional independence assumptions given in Equation~\ref{eq:c1}, we can consider the objective function for each node separately, as only one tie from one node is changing at a time in the CTMC. The node $i$, which is the node that is given the opportunity to change at the current time point, is called the \textit{ego} node. It has the potential to interact with all other nodes in the network, $j \neq i$. These nodes $j$, are referred to as \textit{alter} nodes, and do not change any of their ties when $i$ may change. For the ego node $i$ in the current network state $x$, its \emph{objective function}, which it tries to maximize, is written as
\begin{equation} \label{eq:of1}
f_i(\boldsymbol{\beta}, x) = \sum_k \beta_k s_{ik}(x, \mathbf{Z}),
\end{equation}
for $x \in \mathcal{X}$, the space of all possible directed networks with the $n$ nodes, and $\mathbf{Z}$, the matrix of covariates. The vector $\boldsymbol{\beta}$ contains the parameters of the model with corresponding network and covariate statistics, $s_{ik}(x, \mathbf{Z})$, for $k = 1,\dots, K$. To increase the value of $f_i$, $i$ will destroy one tie where $x_{ij} = 1$, create one tie where $x_{ij}=0$, or make no change if any changes result in lower values of $f_i$.

@@

\subsubsection{Fitting Models to Data}

@@

\section{Example Data and Models}

<<adjmatviz, fig.align='center', fig.cap="A visualization of the adjacency matrices of the three waves of network observations in the ``Teenage Friends and Lifestyle Study'' data. The subset we will be using is outlined in red.", results='hide',fig.show='asis', fig.width=8, fig.height=3, eval = FALSE>>=
source("code/00e_small_friends.R")
view_rs2$wave <- factor(view_rs2$wave)
levels(view_rs2$wave) <- c("Wave 1", "Wave 2", "Wave 3")
ggplot(data= view_rs2, aes(x = y, y=-x)) +
  geom_tile(aes(fill = as.factor(value))) +
  scale_fill_manual("i likes j", labels=c("no", "yes"),
                    values = c('white', 'black')) +
  geom_rect(data= NULL, inherit.aes = FALSE, color = 'red',
            aes(xmin = 19.5, xmax = 35.5, ymax = -19.5, ymin = -35.5, fill =NA)) +
  facet_wrap(~wave) + theme_bw() +
  labs(x = "node j (alter)", y = "node i (ego)") +
  ThemeNoNet +
  theme(aspect.ratio=1, axis.text = element_blank(), axis.ticks = element_blank(), panel.grid = element_blank())
@

<<alldataviz, fig.cap="The smaller friendship network data we will be modelling throughout the paper. The nodes change color as their drinking behavior changes, and the network becomes less densely connected over time. These are the mechanisms we hope to capture with SAOMs.", out.width='.99\\textwidth', fig.height=3>>=
source("code/00e_small_friends.R")
view_rs2$wave <- factor(view_rs2$wave)
levels(view_rs2$wave) <- c("Wave 1", "Wave 2", "Wave 3")
actual1 <- merge(data.frame(as.edgelist(as.network(fd2.w1))),
                 data.frame(id = 1:16, drink = drink2[,1]),
                 by.x = "X1", by.y = "id", all = T)
actual1$Wave <- 1
actual2 <- merge(data.frame(as.edgelist(as.network(fd2.w2))),
                 data.frame(id = 1:16,drink = drink2[,2]),
                 by.x = "X1", by.y = "id", all = T)
actual2$Wave <- 2
actual3 <- merge(data.frame(as.edgelist(as.network(fd2.w3))),
                 data.frame(id = 1:16,drink = drink2[,3]),
                 by.x = "X1", by.y = "id", all = T)
actual3$Wave <- 3
alldat <- rbind(actual1, actual2, actual3)
#alldat$X1 <- paste0("V", alldat$X1)
alldat$X1 <- as.factor(alldat$X1)
#alldat$X2 <- ifelse(!is.na(alldat$X2), paste0("V", alldat$X2), NA)
alldat$X2 <- as.factor(alldat$X2)
#write_csv(alldat, "smallfriends4Geomnet.csv")
alldataviz <- ggplot(data = alldat,
       aes(from_id = X1, to_id = X2, color = as.ordered(drink), group = as.ordered(drink))) +
  geom_net(fiteach = FALSE, directed=TRUE, layout.alg = 'fruchtermanreingold', size = 5, arrowsize = .25, arrowgap = .05, ecolour = 'grey40', labelon=T, labelcolour = 'black', fontsize = 4, vjust = .5, linewidth=0.5) +
  scale_color_brewer(palette = "YlOrRd", name = "Drinking\nBehavior", labels = c("never", "1-2 times per year", "once per month", "once per week")) +
  ThemeNet +
  facet_wrap(~Wave, labeller = "label_both") +
  theme(legend.position = 'bottom',
        panel.background = element_rect(fill = NA, color = "grey30"),
        legend.text = element_text(size = 12, family = "Times New Roman"),
        legend.title = element_text(size = 12, family = "Times New Roman"))
alldataviz
@

To guide our visual exploration of SAOMs, we use two example data sources. The first is a subset of the 50 actor dataset from the ``Teenage Friends and Lifestyle Study" that is provided on the \texttt{RSiena} webpage. These data come from \citet{friendsdata}, and we chose to only work with a subset of the data to make node-link diagrams easier to see and to make any changes in the network more noticeable. %To determine which subset to select, we visualized all waves of the full network using the adjacency matrix visualization approach, which we show in Figure~\ref{fig:adjmatviz}. This adjacency matrix visualization is different from the one in Figure~\ref{fig:toyplots} because it does not show the node IDs on the axes. The network in Figure~\ref{fig:adjmatviz} is much larger than our toy data example, so we remove the node labels to remove clutter. In all three adjacency matrices, the ego and the alter nodes are ordered by node ID, 1-50, which were determined arbitrarily by the relationships in wave 1. 
%In Figure~\ref{fig:adjmatviz}, we show the adjacency matrix visualization of all three observations of this network, with the subset we selected for analysis outlined in red. This subset contains 16 actors, their ties and their drinking behavior for three timepoints. 
The node-link visualization of this data is provided in Figure~\ref{fig:alldataviz}. For model fitting, we condition on wave 1 and estimate the parameters of the models for the second and third waves. The actor-level categorical covariate, drinking behavior, has four values: (1) does not drink, (2) drinks once or twice a year, (3) drinks once a month, and (4) drinks once a week. In Figure~\ref{fig:alldataviz}, the nodes are colored according to the drinking behavior of that student. Over time, we can see that the students tend to drink more and become increasingly isolated into smaller groups. An analysis of this type of data with a SAOM should capture these dynamics in a way that allows the researcher to draw conclusions about the nature of the network and behavioral forces at play.

@@

\subsection{Models}

@@

\section{Model in the data space}\label{sec:minds}

@@

<<summnetM1, message=FALSE, fig.cap='On the left, the first wave of observed data that is conditioned on in the model. On the right, the second wave of observed data. In the middle, a summary network from the first model fit to the data. This summary network represents 1,000 simulations of wave 2 using the values from the simple fitted model M1.',results='hide', fig.height=2.5>>=
friendshipData <- array(c(fd2.w1, fd2.w2, fd2.w3), dim = c(16, 16, 3))
friendshipData <- array(c(fd2.w1, fd2.w2, fd2.w3), dim = c(16, 16, 3))
drink1 <- drink[20:35,1]
drink2 <- drink[20:35,2]
drink3 <- drink[20:35,3]
friendship <- sienaDependent(friendshipData)
alcohol <- varCovar(cbind(drink1,drink2, drink3))
mydata <- sienaDataCreate(friendship, alcohol)
M1eff <- getEffects(mydata)
M2eff <- includeEffects(M1eff, jumpXTransTrip, interaction1 = "alcohol")
M3eff <- includeEffects(M1eff, nbrDist2twice)
simu2 <- read.csv("data/simulation-1000-M1-M2-M3.csv")
means <- simu2 %>% group_by(Model, parameter) %>% summarize(
  means = mean(estimate)
)
M1parms <- (means %>% filter(Model == "M1"))$means
M2parms <- (means %>% filter(Model == "M2"))$means
M3parms <- (means %>% filter(Model == "M3"))$means
set.seed(4231352)
#M1sims1000 <- netvizinf::saom_simulate(dat = mydata,
#                            struct = M1eff,
#                       parms = as.numeric(M1parms),
#                       N = 1000)
#save(M1sims1000, file = "data/M1sims1000.RData")
load("data/M1sims1000.RData")
# M2sims1000 <- netvizinf::saom_simulate(dat = mydata,
#                             struct = M2eff,
#                        parms = as.numeric(M2parms),
#                        N = 1000)
# save(M2sims1000, file = "data/M2sims1000.RData")
load("data/M1sims1000.RData")
# M3sims1000 <- netvizinf::saom_simulate(dat = mydata,
#                             struct = M3eff,
#                        parms = as.numeric(M3parms),
#                        N = 1000)
# save(M3sims1000, file = "data/M3sims1000.RData")
load("data/M3sims1000.RData")

#M1simsdf <- netvizinf::sims_to_df(M1sims1000)
# write_csv(M1simsdf, "data/M1simsdf.csv")
M1simsdf <- read_csv("data/M1simsdf.csv")
#M2simsdf <- netvizinf::sims_to_df(M2sims1000)
#write_csv(M2simsdf, "data/M2simsdf.csv")
M2simsdf <- read_csv("data/M2simsdf.csv")
# M3simsdf <- netvizinf::sims_to_df(M3sims1000)
# write_csv(M3simsdf, "data/M3simsdf.csv")
M3simsdf <- read_csv("data/M3simsdf.csv")

M1avgW2 <- M1simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))
M2avgW2 <- M2simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))
M3avgW2 <- M3simsdf %>% filter(!is.na(to) & wave == 1) %>%
  group_by(from, to) %>%
  summarise(count = n()) %>%
  mutate(weight = ifelse(from == to, 0, count))

@@

For network data, we view an ``average" network through a sort of summary network drawn using the traditional node-link diagram. In Figure~\ref{fig:summnetM1}, we show an \textit{average network} created with 1,000 simulations of the second wave of the network from Model 1. To make this average network, we first simulated 1,000 wave 2 and wave 3 observations of our small friendship example data from model M1, for which parameters had previously been estimated. We then combine the 1,000 simulated instances of wave 2, and count up the number of times each edge appears in a network simulation. Then, we combine these 1,000 networks into a single network with edgeweight equal to the proportion of time that edge appears in our 1,000 simulations. This weighted edgelist is the network we draw in Figure~\ref{fig:summnetM1}. An edge is only drawn in the average network if it appears in at least 51 of the 1000 simulations, with edges that appear with greater frequency emphasized by thicker linewidths and a darker color. On either side of the average network in Figure~\ref{fig:summnetM1}, we show the actual data, wave 1 on the left, and wave 2 on the right. We can see that the structure of the average network is much more similar to the first wave than to the second wave. However, the simulations are supposed to represent the second wave of data on the right of Figure~\ref{fig:summnetM1}. This is an indication that the simple model, M1, is doing a very poor job of capturing the change mechanism from the first to the second wave of observation. The average network can thus be used to help determine model goodness-of-fit. Because the the average network looks more like the first wave than the second wave, we can use the visualization in Figure~\ref{fig:summnetM1} as evidence of poor model fit.

<<summary2, fig.height=2.5, fig.cap="Network differences between the average network from M1 and the first and second data waves. The green edges represent edges in the average network that are not in the data, while red edges are in the data but not in the average network. Black edges are in both the data and the average, and the edgeweight represents the rate at which that edge shows up in simulations.">>=
am_t1 <- as.matrix(igraph::as_adj(igraph::graph_from_data_frame(arrange(t1, parse_number(from), parse_number(to)))))[-17,-17]

diff1 <- am_t1 - fd2.w1
diff1[which(diff1==1)] <- 2
diff1[which(fd2.w1 == 1 & am_t1 == 1)]  <- 1
diff1 <- diff1 %>% dplyr::as_data_frame() 
names(diff1) <- paste0("V", 1:16)
diff1 %>% mutate(from =paste0("V", 1:16)) %>% gather(to, value, -from) %>%
  mutate(from = as.character(parse_number(from)), to = as.character(parse_number(to)), 
         var = recode(value, `0` = "Neither", `2` = "InAvgNotW1",  `1`= "InBoth"),
         color2 = recode(value, `0`="white", `1`= "black", `2` = "green")) %>%
  filter(value !=0) %>% 
  full_join(t1, c("from","to")) -> t1_1 
t1_1 <- arrange(t1_1, from, to)
p1 <- ggplot(data = t1_1) +
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.2, "cm")), ecolour = t1_1$color2,
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 4, arrowgap = .02,
           singletons = T) +
  labs(title = "Difference between Wave 1 and Average from M1") + ThemeNet +
  theme(plot.background = element_rect(color = 'black'))

diff2 <- -am_t1 + fd2.w2 
diff2[which(diff2==1)] <- 2
diff2[which(fd2.w2 == 1 & am_t1 == 1)]  <- 1
diff2 <- dplyr::as_data_frame(diff2)
names(diff2) <- paste0("V", 1:16)
diff2 %>% mutate(from =paste0("V", 1:16)) %>% gather(to, value, -from) %>% 
  mutate(from = as.character(parse_number(from)), to = as.character(parse_number(to)), 
         var = recode(value, `0` = "Neither", `-1` = "InAvgNotW2",  `1`= "InBoth", `2` = "InW2notAvg"),
         color2 = recode(value, `0`="white", `1`= "black", `2` = "red", `-1` = "green")) %>% 
  filter(value !=0) %>% 
  full_join(t1, c("from","to")) -> t1_2 
t1_2$linewidth[is.na(t1_2$linewidth) & !is.na(t1_2$to)] <- 1
t1_2 <- arrange(t1_2, from, to)
t1_2 <- t1_2 %>% filter(!(from == 16 & is.na(to)))
p2 <- ggplot(data = t1_2) +
  geom_net(aes(from_id = from, to_id = to, linewidth= linewidth),
           directed = T, curvature = .2, labelon=T, size = 5,
           arrow = arrow(angle = 20, type = "open", length = unit(.2, "cm")), ecolour = t1_2$color2,
           color = "#252525", hjust = .5, vjust = .5, labelcolour = 'white', fontsize = 4, arrowgap = .02,
           singletons = T) +
  labs(title = "Difference between Wave 2 and Average from M1") + ThemeNet +
  theme(plot.background = element_rect(color = 'black'))
grid.arrange(p1, p2, nrow = 1)
@

In Figure~\ref{fig:summary2}, we show \textit{difference networks} between the average network from M1 and the first and second waves of data. The lack of model fit is very obvious here, with the green edges representing ties that only show up in the average network, and the red edges representing ties in the data that the model misses completely.  Again, the edgeweights demonstrate the frequency of the ties in the simulations from M1. On the left in Figure~\ref{fig:summary2}, we see the difference between the first wave of data and the average network. We only see green and black edges, meaning that the simulated networks consist primarily of edges in wave 1, plus some new edges that aren't in the first wave of data that are reciprocating the ties that are in the data. On the right in Figure~\ref{fig:summary2}, we see the difference between the second wave of data and the average network. Remember, the model is simulating the second wave of data, which is how the average network is created. There are six red edges, which are edges in wave 2 that do not appear at all in the average network. In addition, there are several green edges that are in the average network, but not in wave 2. Finally, there are a few black edges that appear in both. For all edges but the red ones, the linewidth represents the frequency of the edge in the simulations.

@@

% \begin{itemize}
% \item WHat is the model? the network model (SAOM model is just one of many)
% \item What is the data? Edge data, node data, observed time points, unobserved timepoints (CTMC)
% \item Note: specifically dealing with dynamic networks observed at discrete time points.
% \end{itemize}
% What are some things that could go in here? need more data examples; not just the friends data.
% Harry Potter data??? I want to find my own data, too.}

\section{Collections of models}

@@

\subsection{Exploring the space of all possible models}\label{sec:allpossible}

@@

\subsection{Varying model settings}

@@

\subsection{Fitting the same model to different data}\label{sec:samemoddiffdat}

@@

\subsection{Fitting the same model to the same data}\label{sec:samemodsamedat}

@@

\section{Explore algorithms, not just end result}\label{sec:algorithms}
